---
title: "Practical Machine Learning Course Project - Tamás Sallai"
output: html_document
---

## Setting up and loading the data

First, the libraries must be loaded. To greatly speed up the process, muliple cores will be used for the training.

```{r}
library(caret)
library(doMC)
registerDoMC(cores = 4)
```

Then load the data to the variable _data_.

```{r}
data <- read.csv('pml-training.csv')
set.seed(0)
```

After inspecting the data, there are many NA-only columns and some that are non-numerical. Since the gyroscope and the accelerometer data are all numbers, let's filter out these columns. Also as the first 4 are not measurements but metadata, they should be ignored too.

```{r}
filtered <- data[,sapply(data, is.numeric) & colSums(is.na(data))==0]

filtered <- filtered[,-(1:4)]

filtered$classe <- data$classe
```

## Building the model

In order to be able to measure out of sample error rate, let's split the data to a training and a testing set.

```{r}
inTrain <- createDataPartition(y = filtered$classe, p = .7, list = FALSE)

training <- filtered[inTrain,]
testing <- filtered[-inTrain,]
```

The main part is building the model using the training data. In this paper I use the Random Forest approach as it tends to give good results.

```{r}
modFit <- train(classe ~ ., data = training, method = "rf", model = FALSE)
```

## Evaluating the model

After we have a model, let's make a table to get a better picture of it's accuracy.

```{r}
pred <- predict(modFit,testing);
testing$predRight <- pred == testing$classe
table(pred, testing$classe)
```

The in sample error rate:

```{r}
sum(predict(modFit, training) == training$classe) / nrow(training)
```

The out of sample error rate:

```{r}
sum(predict(modFit, testing) == testing$classe) / nrow(testing)
```

## Validation

The validation data set can be downloaded separately.

```{r}
validation <- read.csv('pml-testing.csv')
```

Then the classifications on the data using the previously built model:

```{r}
pred <- predict(modFit,validation)
pred
```